Applying Spatio-temporal Scan Statistics and Spatial Autocorrelation Statistics to identify Covid-19 clusters in the world.

### 2021-01-12 Notebooks/paper_covid.rmd
Update paper_covid.rmd

  In April global cases of Covid-19 reached 1 million, in the same month the world also reached 3 million cases $\!\!^3$, with an exponential growth in the number of infections, the Global Moran's I was also very high for the number of deaths per inhabitant, while local spatial association was strong only in Europe, the boxmap **(Picture 3)** shows that all areas identified with local dependence for the number of deaths per inhabitant were either HH (High-High) clusters in April or HL (High-Low) clusters after April, which means that in general locations with a high number of deaths were close to other areas that also had a high number of deaths. Even though April was a very concerning month to the international community, global spatial autocorrelation on the number of cases per inhabitant only started to grow higher than IM>0.50 in October, but the boxmaps for the number of infections show the same trend that happened for the number of deaths **(Picture 4)**, when looking to the *LISA* map **(Picture 2)**.

>>>>

  In April global cases of Covid-19 reached 1 million, in the same month the world also reached 3 million cases $\!\!^3$, with an exponential growth in the number of infections, the Global Moran's I was also very high for the number of deaths per inhabitant, while local spatial association was strong only in Europe, the boxmap **(Picture 3)** shows that all areas identified with local dependence for the number of deaths per inhabitant were either HH (High-High) clusters in April or HL (High-Low) clusters after April, which means that in general locations with a high number of deaths were close to other areas that also had a high number of deaths. Even though April was a very concerning month to the international community, global spatial autocorrelation on the number of cases per inhabitant only started to grow higher than IM>0.50 in October, but the boxmaps for the number of infections **(Picture 4)** show the same trend that happened for the number of deaths, when looking them together with the *LISA* maps **(Picture 2)** all locations identified with spatial association were either HH or HL *clusters*.

  The Poisson Scan Statistic calculated for the number of cumulative cases identified the duration of emerging *clusters* in the number of infections throughout the year, in a comparison of a timeline of events that occurred in the pandemic $\!\!^3$ and the observed *clusters*, since March 25th the US was identified as an emerging *cluster* by the Scan Statistic and one day later turned out to be the country with most cases of Covid-19 in the World, in May 22th Brazil was identified as a *cluster* and in the same day surpassed Russia turning out to be the 2 $\!\!^n\!^d$ country with most cases in the world. Other patterns can be verified when comparing results given by the Scan Statistics against the locations identified as HH and HL *clusters* in the Boxmaps of infections per inhabitants **(Picture 4)**, all *clusters* of the Scan Statistic were also HH or HL *clusters* in the months they have been identified (October, November and December).
